A Linked Open Social Dataset for Scientific Benchmarking

### 2019-06-30 mdpi/template.tex
adiantada adaptacao p revista data

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{A Linked Open Social Dataset for Scientific Benchmarking}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
The fields of social network analysis and complex networks
are widely researched.
In fact, a myriad of results have been reported which are based in
diverse data, most often not accessible to researchers other than the publishing authors.
In order to furnish the scientific community with a common repertoire,
this work presents an open dataset with diverse provenance of 
human social online networking.
Current data was obtained in 2013-16 from well-known online social networking platforms:
Facebook, Twitter, IRC, Email.
Data from ParticipaBR, AA, and Cidade Democrática are also supplied as samples of less known instances.
The linked data representation was adopted in order to
comply with current best practices, homogenize access, facilitate discovery
and analyzes which integrate third party and provided instances.
Futhermore, the method for obtaing and making the data available is potentially novel, and may of use for other research groups.
This document presents a description, usage remarks and overall statistics of the dataset, which should favor subsequent work in complex and social networks
  and enhancements to the dataset.
}

@@

This work presents a collection of linked and open data
derived from online human social networks.
The dataset is named LOSD (Linked Open Social Data),
and has been achieved by linking social data into a dataset derived from diverse
provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
Such data is proposed as a common and public repertoire for scientific
research involving networks and textual content.
In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant ), and should be used in further research in complex networks and data visualization (FAPESP grant ) which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.

\subsection{Related work}
Most available datasets on human social networks rely on only one source~\cite{o1,o2,o3,o4},
on specialized communities such as related to research and movies~\cite{s1,s2,s3,s4,s5},
or present only statistics about social networks, and not the data that yields the networks~\cite{st1,st2,st3,st4}.
Contributions which are most similar to LOSD are mentioned in this section.
In~\cite{fb1}, data obtained only from Facebook is reported.
With such a focus, the authors made available a number of statistics,
but the data is offline at least since 2010.
Linked human social network data is scarcely found in the scientific community.
In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud, and made available a dataset from the Advogato platform,
dedicated to free software developers linked by trust relatioships.
There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2,t3,t4}),
most often with a specific focus (users, tweets, hashtags, etc)
and not presented to the scientific community by means of a research article.
In summary, no human social network dataset presented to the
scientific community containts the diverse provenance used in LOSD,
potentially because researchers did not have the means to acquire and
share the data as described in Section~\ref{acq}.

\section{Data Description}\label{materials}
\subsection{Data outline}
\input{../misc/overallText}

\input{../tables/nsnapshots}
\label{outline}
Current overall results concern data selection and preparation for data discovery.
The main result is the data made available, which enables benchmarking of scientific results
and easy experimentations, and no compliant dataset was found by the authors.
Secondary results include data outline through figures and tables,
software support and example SparQL queries.

Data was gathered from:
\begin{itemize}
    \item public APIs (Twitter, Email);
    \item public logs (IRC and AA);
    \item Netvizz software~\cite{netvizz} and subsequent donation by users (Facebook);
    \item donation by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
\end{itemize}

Integration and uniformity of access is obtained through linked data
representation, as described in Section~\ref{queries}.
This section introduces the underlying data in very concise terms.
One should access the Supporting Information document for
a more information about the dataset.

\subsection{Snapshots}
Of central importance to presented database is the concept of a snapshot.
A snapshot is herein a set of data gathered together, at a contiguous time
unit.
Examples: the first 20 thousand email messages of an email list
comprises a snapshot; the tweets from the MAMA music event is a
snapshot; the friendship, interaction and posts structures of a facebook
group, prospected at the same time, is a snapshot.

\subsection{Facebook data}
Friendship ego networks (networks whose nodes represent friends of a user)
were donated from individual users in 2013 and 2014.
Friendship and interaction networks from groups were gathered from
groups where the first author was a participant.
Additionally, some groups have post texts along some metadata, such as
the number of likes.

\subsection{Twitter data}
Tweets were gathered through the Twitter streaming public API.
Each snapshot is unified by a distinct hashtag.
Network links are canonically yield by retweets,
but replies and user mentions are also kept in the database.

\subsection{IRC data}
Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
The database has records of relations yield by directed messages and
mentions.

\subsection{Email data}
Email snapshots refer to individual email lists.
All messages were obtained from the Gmane public email archive~\cite{gmane}.
Each message has the original text and the text without some of the lines
from previous messages or that are software code or author signature.
Most importantly, each message instance holds the ID of the message it is
a reply to, if any, which enables the achievement of interaction networks~\cite{bird,stab}

\subsection{ParticipaBR data}
ParticipaBR was a Brazilian federal platform for social participation,
the most important platform from the myriad of social participation platforms
that were available in 2011-2018~\cite{spbr}.
Texts are derived from blog posts and networks may be obtained from
friendship and interaction criteria.

\subsection{Cidade Democr\'atica data}
Cidade Democrática is a Brazilian civil society social participation portal.
Data gathered is complex in the number of types of instances.

\subsection{AA data}
AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications,
it is based on testifying and sharing ongoing work.
The data was gathered from different versions of the system and from an IRC
log.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
LOSD is expressed as linked data through RDF
and is ontologicaly described through the
data-driven ontology synthesis method presented in Section~\ref{ont}.

\subsection{Data acquisition}\label{acq}
Research involving human subjects raise well-known ethic issues~\cite{}.
Furthermore, the online social networking platforms often restrict access
to the data and declare illegal the use of their data through web crowlers~\cite{ile1,ile2}.
In this context, research using data, e.g. from Facebook, is considered
problematic if not prohibited.
We reached a potential solution to this issue through continuated research with
collaborators in the fields of phylosophy, social sciences, and psychology.
The conceptual framework is presented in~\cite{antphy,antphy2}, and in essence
consists in researching first the social networks related to the researchers,
and keeping the procedures, data, algorithms and results publicly available.
This practices enabled collecting and sharing the data in LOSD.
Notice that, although Twitter data is available through public APIs, the other sources are more problematic.

% download via netviz when fb allowed it, donations
% twitter, bots, gmane
% contact with developers of AA, PBR, CD
% Anthropological physics

\subsection{Linked open data}
Linked data refers to data published in the web in such a way that it is
machine readable and conforms to a set of best practices.
The yielded web of data is composed of documents on the web
such as the web of HTML documents.
In practice, the publication of linked data can me summarized
by 1) the use of RDF to publish data on the web and 2) the use of RDF
links to interlink data from different sources~\cite{ld1,ld2}.
The web is expected to be interconnected and to grow by the systematic application of four
steps~\cite{lee1}:
\begin{itemize}
    \item Use URIs to identify things~\cite{uri}.
    \item Use HTTP URIs.
    \item Provide useful information when a URI is accessed via HTTP.
    \item Provide other URIs in the description of resources so human
        and machine agents can perform discovery.
\end{itemize}

The Linked Open Data~\cite{lod} is a constantly growing cloud of data,
the global data space, which is usually
conceived as centered around the DBPedia, a linked data representation
of data from Wikipedia~\cite{dbpedia0,dbpedia},
and sometimes called the Giant Global Graph (GGG, akin to the WWW).

@@

\subsection{Data-driven ontology synthesis}\label{ont}
OWL Ontologies are tools to describe taxonomies and the
structure of knowledge~\cite{owl1}.
Most ontologies are created by domain experts even though the data they
arrange is often given by a software system and has a predefined
structure~\cite{owl2}.

@@

Supporting Information document of this article.

@@

\section{Usage notes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt} 
\subsection{Software tools}
The authors made available software for rendering analyses and
the data set itself.
\subsubsection{Triplification routines}
For each social platform there is a \emph{triplification} routine,
i.e. a script for expressinf the data as RDF.
Original formats and further observations are presented in
Table~\ref{tab:provenance}.
  {\renewcommand{\arraystretch}{1.2}
\begin{table*}[h!]\scriptsize
\begin{center}
\caption{Social platforms, original formats and further observations for
the database.}\label{tab:provenance}
\begin{tabular}{ l | p{3cm} p{3cm} c }
    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\specialrule{1.5pt}{1pt}{1pt}
    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
    Email & mbox & obtained through the Gmane public archive & Gmane~\cite{gmane} \\\hline
    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social~\cite{social} \\\hline
    IRC & plain text log & obtained through Supybot logging & Social \\\hline
    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
    Twitter & JSON & obtained through Twitter streaming API & Social \\
\end{tabular}\end{center}
\end{table*}                    
  }
\subsubsection{Topological and textual analysis}\label{ana}
Routines are available for taking topological and textual measures from
the LOSD data.
  Auxiliary routines, such as for performing principal component analysis (PCA)
  and obtaining statistics derived from the Kolmogorov-Smirnov test~\cite{kol}, are available to ease pattern recognition.

>>>>

%=================================================================
\documentclass[data,datadescriptor,submit,moreauthors,pdftex]{Definitions/mdpi}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{LOSD: A Linked Open Social Dataset for Scientific Benchmarking}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
The fields of social network analysis and complex networks
are widely researched.
In fact, a myriad of results have been reported which are based in
diverse data, most often not accessible to researchers other than the publishing authors.
In order to provide the scientific community a common repertoire,
this work presents an open dataset with diverse provenance of 
human social online networking.
Current data was obtained in 2013-16 from well-known online social networking platforms:
Facebook, Twitter, IRC, Email.
Data from ParticipaBR, AA, and Cidade Democrática are also supplied as samples of less known instances.
The linked data representation was adopted in order to
comply with current best practices, homogenize access, facilitate discovery
and analyzes which integrate third party and provided data.
Futhermore, the method for obtaing and making the data available is potentially novel,
and may be of use for other research groups.
This document presents a description, usage remarks and overall statistics of the dataset,
which should favor subsequent work in complex and social networks and enhancements to the dataset.
}

@@

This work presents a linked and open dataset
derived from online human social networks.
The dataset is named LOSD (Linked Open Social Data),
and has been achieved by linking social data into a dataset derived from diverse
provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
Such data may be of use as a common and public repertoire for scientific
research involving complex networks, specially if social networks and textual content are pinpointed.
In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant ),
and should be further used e.g. in researching complex networks and data visualization (FAPESP grant )
which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.

Data collection was performed by a potentially innovative method which places the researcher
at the core of data collection, and keeps the resources (e.g. code, data, documents) open,
in order to ammeliorate ethical issues involved in studying human social systems.

This paper is organized as follows. Next subsection briefly discusses related work.
Section~\ref{materials} describes the data currently available.
Section~\ref{methods} describes the methods used for data gathering, translation into RDF linked data, and the synthesis of auxiliary structures.
Section~\ref{usage} contains direction for using the data.

\subsection{Related work}
Most available datasets on human social networks rely on only one source~\cite{o1,o2,o3,o4},
on specialized communities such as related to research or movies~\cite{s1,s2,s3,s4,s5},
or present only statistics about social networks, and not the data that yields the networks~\cite{st1,st2,st3,st4}.
Contributions which are most similar to LOSD are mentioned in this section.
In~\cite{fb1}, data obtained only from Facebook is reported.
With such a focus, the authors made available a number of statistics,
but the data is offline at least since 2010.
Linked human social network data is scarcely found in the scientific community.
In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud, and made available a dataset from the Advogato platform,
dedicated to free software developers linked by trust relatioships.
There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2,t3,t4}),
most often with a specific focus (users, retweets, hashtags, etc)
and not presented to the scientific community by means of a research article.
In summary, no human social network dataset presented to the
scientific community containts the diverse provenance used in LOSD,
potentially because researchers did not have the means to acquire and
share the data as described in Section~\ref{acq}.

\section{Data Description}\label{materials}
\subsection{Data outline}\label{outline}
% scripts/percolationUse2.py
The database consists of 31,996,740 triples, 3,030,433 network links yield by main interactions or relations, 350,284 participants and 245,194,377 characters. Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.
Appendix~\ref{aoverall} holds the queries used to obtain such data.
This section describes de data in LOSD in very brief terms.
The Supplementary document is provided for further information about the dataset.

\noindent Data was gathered from:
\begin{itemize}
    \item public APIs (Twitter, Email);
    \item public logs (IRC and AA);
    \item Netvizz software~\cite{netvizz} and subsequent donation by users (Facebook);
    \item donation by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
\end{itemize}

\subsection{Snapshots}
Of central importance to the presented dataset is the concept of a snapshot.
A snapshot is herein a set of data gathered together, at a contiguous time
unit.
Examples: the first 20 thousand email messages of an email list
comprises a snapshot; the tweets from the MAMA music event is a
snapshot; the friendship, interaction and posts structures of a facebook
group, prospected at the same time, is a snapshot.
Table~\ref{tsnap} holds the number of snapshots from each provenance in LOSD.

\input{../tables/nsnapshots}

\subsection{Facebook data}
Friendship ego networks (networks whose nodes represent friends of a user and links represent friendships)
were donated from individual users in 2013 and 2014.
Friendship and interaction networks from groups were gathered from
groups where the first author was a participant.
Additionally, some groups have post texts along some metadata, such as
the number of likes.

\subsection{Twitter data}
Tweets were gathered through the Twitter streaming public API.
Each snapshot is unified by a distinct hashtag.
Network links are canonically yield by retweets,
but replies and user mentions are also kept in the database.
LOSD kept the main data fields returned by Twitter API, such as creation date, text, and hashtags.

\subsection{IRC data}
Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
LOSD holds records of relations yield by directed messages and
mentions and the text related to each message.

\subsection{Email data}
Email snapshots refer to individual email lists.
All messages were obtained from the Gmane public email archive~\cite{gmane}.
Each message has the original text and the text without some of the lines
from previous messages or that are software code or author signature.
Most importantly, each message instance holds the ID of the message it is
a reply to, if any, which enables the achievement of canonical interaction networks~\cite{bird,stab}.

\subsection{ParticipaBR data}
ParticipaBR is a Brazilian federal platform for social participation,
the most important platform from the myriad of social participation platforms
that were available in 2011-2018~\cite{spbr}.
Texts are derived from blog posts and networks may be obtained from
both friendship and interaction criteria.

\subsection{Cidade Democrática data}
Cidade Democrática is a Brazilian civil society social participation portal.
Data gathered is complex in the number of types of instances as noticeable in the Supplementary document.

\subsection{AA data}
AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications by sharing sentences about ongoing work and testifying about their quality.
The data was gathered from different versions of the system and from an IRC
log.

\section{Methods}\label{methods}
LOSD is expressed as linked data through RDF
and is ontologicaly described through the
data-driven ontology synthesis method presented in Section~\ref{ont}.

\subsection{Data acquisition}\label{acq}
Research involving human subjects raise well-known ethic issues~\cite{eth}.
Furthermore, the online social networking platforms often restrict access
to the data and declare illegal the use of their data through web crowlers~\cite{ile1,ile2}.
Although Twitter and GMANE (email) data is available through public APIs,
the other sources are more problematic.
In this context, research using data, e.g. from Facebook, is precarious
if not prohibited.
We reached a potential solution to this issue through continuated research with
collaborators in the fields of phylosophy, social sciences, antropology and psychology.
The conceptual framework is presented in~\cite{antphy,antphy2}, and in essence
consists in researching first the social networks related to the researchers,
and keeping the procedures, data, algorithms and results publicly available.
These practices enabled collecting and sharing the data in LOSD,
and was called ``Anthropological Physics'' because it relies in the
etnographic tecnique of investigating the social structures
by studying the researcher, as to ammeliorate the ethic issues.
Accordingly, aside from the data acquired from Twitter, all data in LOSD
was gathered from social instances where the first author is a participant.
For example, all Facebook ego networks were donated to the first author by
users which know him and were interested in his research and intervention on the networks,
and all email and Facebook groups data was obtained from groups where the first author is a member.

Facebook data was obtained through the Netvizz software~\cite{netviz},
Facebook allowed downloading friendships and interaction data as available in LOSD.
Data from Twitter was obtained through the Twitter streamming API.
Data From IRC was obtained through bots that kept logs of the messages.
Email data was obtained through the GMANE public API.
Data from ParticipaBR, Cidade Democrática, and AA was obtained through contact
of the researcher with the developers of the platforms.

\subsection{Linked open data}
Linked data refers to data published in the web in such a way that it is
machine readable and conforms to a set of best practices.
The yielded web of data is composed of documents on the web
such as the web of HTML documents.
In practice, the publication of linked data can me summarized
by 1) the use of RDF and 2) the use of RDF
links to interlink data from different sources~\cite{ld1,ld2}.
The web is expected to be interconnected and to
grow by the systematic application of four
steps~\cite{lee1}:
\begin{itemize}
    \item Use URIs to identify things~\cite{uri}.
    \item Use HTTP URIs.
    \item Provide useful information when a URI is accessed via HTTP.
    \item Provide other URIs in the description of resources so human
        and machine agents can perform discovery.
\end{itemize}

The Linked Open Data~\cite{lod} cloud is a constantly growing cloud of data,
the global data space, which is usually
conceived as centered around the DBPedia, a linked data representation
of data from Wikipedia~\cite{dbpedia0,dbpedia},
and sometimes called the Giant Global Graph (GGG, akin to the WWW).

@@

\subsection{Data-driven ontology synthesis}\label{ont}
Most ontologies are created by domain experts even though the data they
arrange is often given by a software system and has a predefined
structure~\cite{owl1,owl2}.  

@@

Supplementary document of this article.

@@

\section{Usage notes}\label{usage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt} 
\subsection{Software tools}
The authors made available software for rendering analyses and
the data set itself.
\subsubsection{Triplification routines}
For each social platform there is a \emph{triplification} routine,
i.e. a script for expressinf the data as RDF.
Original formats and further observations are presented in
Table~\ref{tab:provenance}.
  {\renewcommand{\arraystretch}{1.2}
\begin{table*}[h!]\scriptsize
\begin{center}
\caption{Social platforms, original formats and further observations for
the database.}\label{tab:provenance}
\begin{tabular}{ l | p{3cm} p{3cm} c }
    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\specialrule{1.5pt}{1pt}{1pt}
    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
    Email & mbox & obtained through the Gmane public archive & Gmane~\cite{gmane} \\\hline
    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social~\cite{social} \\\hline
    IRC & plain text log & obtained through Supybot logging & Social \\\hline
    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
    Twitter & JSON & obtained through Twitter streaming API & Social \\
\end{tabular}\end{center}
\end{table*}                    
  }
\subsubsection{Topological and textual analysis}\label{ana}
Routines are available for taking topological and textual measures from
the LOSD data.
  Auxiliary routines, such as for performing principal component analysis (PCA)
  and obtaining statistics derived from the Kolmogorov-Smirnov test~\cite{kol}, are available to ease pattern recognition.

### 2019-06-30 misc/overallText.tex
adiantada adaptacao p revista data

>>>>

# The database consists of 31,996,740 triples, 3,030,433 network links yield by main interactions or relations, 350,284 participants and 245,194,377 characters. Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.
